---
title: "Cultural Data Science - Exam Project"
subtitle: 'Health Behaviour in School-aged Children'
author: "Laura Wulff Paaby & Rikke Uldbæk"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    df_print: paged
  blogdown::html_page:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#loading packages
pacman::p_load(ggplot2, tidyverse, lme4)

setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Cognitive Science/5th semester/CDS_EXAM/CultDat_EXAM")
```


## Loading Data

The data is loaded from the cloud service Google Drive. 
```{r}
# Loading files from google drive 
id_2014 <- "1Nes6x0xB-M2tnPu4DA0AJbD0hVuiXDc4" # google file ID - hbsc 2014
hbsc2014<- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id_2014)) # data hbsc 2014
```


## Data Preprocessing 
In this section, appropriate filtering and pre-processing of the data will be carried out. 


### Countries of Interest
The countries of interest in this paper are: Denmark, Italy, Spain, England, Latvia. In the hbsc2014 dataframe, all countries are denoted with a country-number, however this is rather confusing, thus we rename the country-number in accordance with the actual country-name (Denmark = 208000, Italy = 380000, Spain = 724000, England = 826001, Latvia = 428000)  

```{r}
# Filtering the five countries 
dat2014 <- hbsc2014 %>% 
   filter(COUNTRYno == 208000 | COUNTRYno == 380000 | COUNTRYno == 724000 | COUNTRYno == 826001 | COUNTRYno == 428000 ) %>% 
  rename(countryno = COUNTRYno,
         surveyyear = year,
         uniqueid = UniqueID) # renaming it to match the others 

# a little data was collected in 2013, this is renamed to 2014 for further analysis
dat2014$surveyyear <- ifelse(dat2014$surveyyear== 2013, 2014, 2014)# making all data from 2013 into 2014
```
Apparently England had some data from 2013 as the only country, we changed this data to being 2014


```{r}
# renaming countries in a function
rename_func <- function(data){
data$country_name <- (ifelse(data$countryno == "208000","Denmark", ifelse(data$countryno == "380000", "Italy", ifelse(data$countryno == "724000","Spain", ifelse(data$countryno == "826001","England", ifelse(data$countryno == "428000","Latvia", ifelse(data$countryno == "80400", "Ukraine", NA)))))))
return(data)                                                                
}

#using the function and checking data 
# dat2014
dat2014 <- rename_func(dat2014)
unique(dat2014$country_name)
```



### Further Subsetting
First we will subset the data of interest for model building.
```{r}
#Check we only have data from 2014
unique(hbsc2014$HBSC)

#subset the important variables 
subset_df2014 <- dat2014 %>% 
  select(uniqueid,AGE,sex, #general variables
         lifesat,#mental health variable
         breakfastwd, breakfastwe, fruits, toothbr,sweets, vegetables, ondiet# nutritional variables
         )

```


### Replace NAs
Then we will identify and replace NA's with the mean
```{r}
#Replace NA's with mean (note makes everything numeric)
subset_df2014[subset_df2014==-99] <- NA #some rows may contain -99 instead of NA (replace then)
subset_df2014[subset_df2014=="NaN"] <- NA #some rows may contain NaN instead of NA (replace then)

subset_df2014 <- subset_df2014 %>% 
  mutate_if(is.numeric, ~replace_na(.,mean(., na.rm = TRUE))) %>% 
  round(digits = 0)
```

Finally country-name is added to the dataframe since it disturbed previous work by being a character class
```{r}
#Adding country to the dataframe 
subset_df2014 <- cbind(subset_df2014, country= dat2014$country_name)

#making country a factor
subset_df2014$country <- as.factor(subset_df2014$country)
```


### Plotting Countries
```{r}
country_plot<- ggplot(subset_df2014, aes(x= country, y=lifesat, colour =country)) + geom_point(stat="summary", fun.y="mean") + 
  geom_errorbar(aes(width = 0.1), stat="summary", fun.data="mean_se", fun.args = list(mult = 10))+ggtitle("")+
  stat_summary(fun.y=mean, aes(group=1), geom="line", colour="grey") +
  stat_summary(fun.y=mean, aes(group=1), geom="point", colour="grey", size=1)+
  labs(x = "Countries", y = "Life Satisfaction") + theme(text=element_text(size=13, family="Georgia"))
  
#rename legend
update_labels(country_plot, list(colour="Country"))
```



## Analysis
In this section we will model nutritional factors and mental health in order to investigate their relationship.

### Models
Various of models are constructed to model nutritional factors on mental health in order to find the best fitting model.
```{r}
## mental health ~ nutritional factors

#####country as random intercept
MH_m1 <- lmer(lifesat ~ breakfastwd + (1 | country), subset_df2014)
MH_m2 <- lmer(lifesat ~ breakfastwd+ breakfastwe+ (1 | country), subset_df2014)
MH_m3 <- lmer(lifesat ~ breakfastwd+ breakfastwe+ fruits+ (1 | country), subset_df2014)
MH_m4 <- lmer(lifesat ~ breakfastwd+ breakfastwe+ fruits+ ondiet +(1 | country), subset_df2014)
MH_m5 <- lmer(lifesat ~ breakfastwd+ breakfastwe+ fruits+ ondiet + sweets + (1 | country), subset_df2014)
MH_m6 <- lmer(lifesat ~ breakfastwd+ breakfastwe+ fruits+ ondiet + sweets + vegetables + (1 | country), subset_df2014)


######country as fixed effects
MH_m7 <- lm(lifesat ~ breakfastwd + country, subset_df2014)
MH_m8 <- lm(lifesat ~ breakfastwd+ breakfastwe+ country, subset_df2014)
MH_m9 <- lm(lifesat ~ breakfastwd+ breakfastwe+ fruits+ country, subset_df2014)
MH_m10 <- lm(lifesat ~ breakfastwd+ breakfastwe+ fruits+ ondiet +country, subset_df2014)
MH_m11 <- lm(lifesat ~ breakfastwd+ breakfastwe+ fruits+ ondiet + sweets + country, subset_df2014)
MH_m12 <- lm(lifesat ~ breakfastwd+ breakfastwe+ fruits+ ondiet + sweets + vegetables + country, subset_df2014)

######interactions
MH_m13 <- lm(lifesat ~ fruits:vegetables + country, subset_df2014)
MH_m14 <- lm(lifesat ~ ondiet:vegetables + country, subset_df2014)
MH_m15 <- lm(lifesat ~ sweets:vegetables + country, subset_df2014)
MH_m16 <- lm(lifesat ~ breakfastwd:vegetables + country, subset_df2014)
MH_m17 <- lm(lifesat ~ breakfastwe:vegetables + country, subset_df2014)
MH_m18 <- lm(lifesat ~ ondiet:fruits + country, subset_df2014)
MH_m19 <- lm(lifesat ~ ondiet:sweets + country, subset_df2014)
MH_m20 <- lm(lifesat ~ ondiet:breakfastwd + country, subset_df2014)
MH_m21 <- lm(lifesat ~ ondiet:breakfastwe + country, subset_df2014)
MH_m22 <- lm(lifesat ~ sweets:fruits + country, subset_df2014)
MH_m23 <- lm(lifesat ~ sweets:breakfastwd + country, subset_df2014)
MH_m24 <- lm(lifesat ~ sweets:breakfastwe + country, subset_df2014)
MH_m25 <- lm(lifesat ~ breakfastwd:fruits + country, subset_df2014)
MH_m26 <- lm(lifesat ~ breakfastwe:fruits + country, subset_df2014)
MH_m27 <- lm(lifesat ~ breakfastwd:breakfastwe + country, subset_df2014)
MH_m28 <- lm(lifesat ~  country:sweets, subset_df2014)
MH_m29 <- lm(lifesat ~  country:fruits, subset_df2014)
MH_m30 <- lm(lifesat ~  country:vegetables, subset_df2014)
MH_m31 <- lm(lifesat ~  country:ondiet, subset_df2014)
MH_m32 <- lm(lifesat ~  country:breakfastwd, subset_df2014)
MH_m33 <- lm(lifesat ~  country:breakfastwe, subset_df2014)
```

### Model Comparison

In this section we will compare the best fitting models. 
```{r}
anova(MH_m1, MH_m2,MH_m3, MH_m4, MH_m5, MH_m6, MH_m7, MH_m8,MH_m9,MH_m10, MH_m11, MH_m12, MH_m13, MH_m14, MH_m15, MH_m16, MH_m17, MH_m18, MH_m19, MH_m20, MH_m21, MH_m22, MH_m23, MH_m24, MH_m25, MH_m26, MH_m27, MH_m28, MH_m29, MH_m30, MH_m31, MH_m32, MH_m33)
```

Evaluating the mental health model using AIC and BIC
```{r}
#writing AIC values down
MH_aic <- AIC(MH_m1, MH_m2,MH_m3, MH_m4, MH_m5, MH_m6, MH_m7, MH_m8,MH_m9,MH_m10, MH_m11, MH_m12, MH_m13, MH_m14, MH_m15, MH_m16, MH_m17, MH_m18, MH_m19, MH_m20, MH_m21, MH_m22, MH_m23, MH_m24, MH_m25, MH_m26, MH_m27, MH_m28, MH_m29, MH_m30, MH_m31, MH_m32, MH_m33)


#writing down BIC values
MH_bic <- BIC(MH_m1, MH_m2,MH_m3, MH_m4, MH_m5, MH_m6, MH_m7, MH_m8,MH_m9,MH_m10, MH_m11, MH_m12, MH_m13, MH_m14, MH_m15, MH_m16, MH_m17, MH_m18, MH_m19, MH_m20, MH_m21, MH_m22, MH_m23, MH_m24, MH_m25, MH_m26, MH_m27, MH_m28, MH_m29, MH_m30, MH_m31, MH_m32, MH_m33)


#putting results into one dataframe:
MH_df <- tibble(Model = c("MH_m1", "MH_m2","MH_m3", "MH_m4", "MH_m5", "MH_m6", "MH_m7", "MH_m8","MH_m9","MH_m10", "MH_m11", "MH_m12", "MH_m13", "MH_m14", "MH_m15", "MH_m16", "MH_m17", "MH_m18", "MH_m19", "MH_m20", "MH_m21", "MH_m22", "MH_m23", "MH_m24", "MH_m25", "MH_m26", "MH_m27", "MH_m28", "MH_m29", "MH_m30", "MH_m31", "MH_m32", "MH_m33"), AIC=MH_aic$AIC, BIC=MH_bic$BIC)

view(MH_df) #Best model 
```

The best model that performs the best is Model 12 (MH_m12) with a BIC = 120043.8 and AIC = 119944.1, p < .001. However, Model 6 (MH_m6) makes conceptually better sense as we allow each country to vary sense, likewise does it only perform minimally worse than Model 12 with a BIC= 120085.8 and AIC = 120011.0, p < .001.

### Checking assumptions
We will check the following assumptions for Multiple Regression
  \- Linarity
  \- Homoschedasticity 
  \- Normality of residuals

__Linearity__ 
It’s called “linear model” for a reason! 
If data is NOT linear, a residual plot will indicate some kind of curve, or it will indicate some other pattern (e.g., two lines if you have categorical binary data).
  -no patterns or curves  PLEASE!!!
```{r}
#residual plot 
plot(MH_m6)
```

__Homoschedasticity__
It says that the variability of your data should be approximately equal across the range of your predicted values (we want equal variances). If homoscedasticity is violated, you end up with heteroskedasticity, or, in other words, a problem with unequal variances.
- look at the plot "residuals VS fitted"
    - WE WANT A CLOUD OF DOTS!

```{r}
#make residual plot
plot(MH_m6)
```


__Normality of residuals__
.. the least important.. it turns out that linear models are relatively robust against violations of the assumptions of normality. 

How to check it? 
- histograms and qq-plot of model
```{r}
#histogram of model
hist(residuals(MH_m6)) #bell shaped =normal

#checking numerically
residuals <- resid(MH_m6) #should have a mean around 0
summary(residuals)

# Q-Q plot of model
qqnorm(residuals(MH_m6)) #dots fall in a line = normal

```

## Results

### Results of the Model

```{r}
#printing summary of the model
summary(MH_m6)
parameters::p_value(MH_m6)
jtools::summ(MH_m6)
```


### Cross Country Stats

Calculating mean and standard deviation of life satisfaction for each country
```{r}
#Denmark 
mean(subset_df2014[subset_df2014$country == 'Denmark', 'lifesat'])
sd(subset_df2014[subset_df2014$country == 'Denmark', 'lifesat'])

#England
mean(subset_df2014[subset_df2014$country == 'England', 'lifesat'])
sd(subset_df2014[subset_df2014$country == 'England', 'lifesat'])


#Italy
mean(subset_df2014[subset_df2014$country == 'Italy', 'lifesat'])
sd(subset_df2014[subset_df2014$country == 'Italy', 'lifesat'])

#Latvia
mean(subset_df2014[subset_df2014$country == 'Latvia', 'lifesat'])
sd(subset_df2014[subset_df2014$country == 'Latvia', 'lifesat'])

#Spain
mean(subset_df2014[subset_df2014$country == 'Spain', 'lifesat'])
sd(subset_df2014[subset_df2014$country == 'Spain', 'lifesat'])
```


